---
title: "Pre-analysis script for data preparation"
author: "Bastien BENNETOT"
date: "13 avril 2017"
output:
  html_document: default
  pdf_document: default
---

#Mushroom observer
## Initialisation
```{r,echo=TRUE}
setwd("~/Bureau/STAGE M1 IMPORTANT/algo")

library(sp)
library(SDMTools)
library(raster)
library(foreach)
library(Hmisc)
library(rgdal)
library(RSQLite)



`%ni%` <- Negate(`%in%`) 

###### Gather observation data
mushroom_observer <- read.table("mushroom_observer_170203.csv", sep=",")
mushroom_observer_all <- read.table("mushroom_observer_all_170220.csv", sep=",")
inaturalist_all <-read.csv( file = "iNaturalist_all_170505.csv")

#data_gbif<-gbif(geo = T,genus = "Amanita")
#write.csv(data_gbif,file = "gbif_amanita_170519.csv")
gbif_amanita<-read.csv("gbif_amanita_170519.csv")
```

## Preparation of the dataframe for amanita only from mushroom observer
```{r,echo=TRUE}
#Add  columns for latitude and longitude coordinate and color
mushroom_observer <- cbind(mushroom_observer,c(0),c(0),c(0),c(0))
#Name columns
colnames(mushroom_observer) <- c("observation", "specname", "north", "west", "east", "south","hemisphere","latitude","longitude","carp_color","spore_color")
#Upress rows without species name
mushroom_observer<-subset(mushroom_observer,specname!="")
#Order by species name the dataframe
mushroom_observer <- mushroom_observer[order(mushroom_observer[,2]), ]

## Longitude and latitude computing from coordinate gathered by php code (north, west, east, south, hemisphere)
#Calculating latitude and longitude based on location coordinate
mushroom_observer$latitude<-(mushroom_observer$north+mushroom_observer$south)/2
mushroom_observer$longitude<-(mushroom_observer$east+mushroom_observer$west)/2

mushroom_observer$latitude[mushroom_observer$hemisphere=="ES"] <- -mushroom_observer$latitude[mushroom_observer$hemisphere=="ES"]
mushroom_observer$latitude[mushroom_observer$hemisphere=="SE"] <- -mushroom_observer$latitude[mushroom_observer$hemisphere=="SE"]
mushroom_observer$longitude[mushroom_observer$hemisphere=="NW"] <- -mushroom_observer$longitude[mushroom_observer$hemisphere=="NW"]
mushroom_observer$longitude[mushroom_observer$hemisphere=="WN"] <- -mushroom_observer$longitude[mushroom_observer$hemisphere=="WN"]

mushroom_observer$latitude[mushroom_observer$hemisphere=="SW"] <- -mushroom_observer$latitude[mushroom_observer$hemisphere=="SW"]
mushroom_observer$latitude[mushroom_observer$hemisphere=="WS"] <- -mushroom_observer$latitude[mushroom_observer$hemisphere=="WS"]
mushroom_observer$longitude[mushroom_observer$hemisphere=="SW"] <- -mushroom_observer$longitude[mushroom_observer$hemisphere=="SW"]
mushroom_observer$longitude[mushroom_observer$hemisphere=="WS"] <- -mushroom_observer$longitude[mushroom_observer$hemisphere=="WS"]

mushroom_observer$longitude[mushroom_observer$hemisphere=="NN"] <- -mushroom_observer$longitude[mushroom_observer$hemisphere=="NN"]

##Correction on mushroom observer localization
#Observation are localized by rectangular zone. Correction with a threshold of height and width. It correspond to an incertitude of 0.1° of latitude and longitude 

mushroom_observer<-subset(mushroom_observer, abs(mushroom_observer$west-mushroom_observer$east)<0.1)
mushroom_observer<-subset(mushroom_observer, abs(mushroom_observer$north-mushroom_observer$south)<0.1)

##Remove some error of mushroom user localization
#When there is no localization sometime it's logitude=-179
  mushroom_observer<-subset(mushroom_observer, mushroom_observer$longitude!=-179.0000)
  mushroom_observer<-subset(mushroom_observer, mushroom_observer$longitude!=179.0000) 

```

#Amanita dataset for color identification
##Link species name to color
Species that are not linked to color are removed
```{r,echo=TRUE}
#Reference dataframe for species to color
species_to_color <- read.table("species_to_color.csv", sep=",", header=TRUE)
species_to_color<-subset(species_to_color, carp_col!="")
species_to_color<-unique(species_to_color)
species_to_color <- species_to_color[order(species_to_color[,1]), ]
species_to_color$carp_col<-as.character(species_to_color$carp_col)
species_to_color$spor_col<-as.character(species_to_color$spor_col)

#Remove undefined species from data
mushroom_observer_color<- mushroom_observer[mushroom_observer$specname%in%species_to_color$spec,]
mushroom_observer_color<-droplevels(mushroom_observer_color)
#Remove duplicate from reference dataframe (species_to_color)
species_to_color<- species_to_color[species_to_color$spec%in%mushroom_observer_color$specname,]
species_to_color<-droplevels(species_to_color)

#Attribution of color to species in data
for(i in 1:nrow(mushroom_observer_color)){
mushroom_observer_color$carp_color[i]<-species_to_color$carp_col[mushroom_observer_color$specname[i]==species_to_color$spec]
mushroom_observer_color$spore_color[i]<-species_to_color$spor_col[mushroom_observer_color$specname[i]==species_to_color$spec]
}

mushroom_observer_color <- subset(mushroom_observer_color, carp_color!="" )
mushroom_observer_color<-droplevels(mushroom_observer_color)
mushroom_observer_color$carp_color<-as.factor(mushroom_observer_color$carp_color)

```

## Preparation of the dataframe for all organisms from mushroom observer
```{r,echo=TRUE}
#Add  columns for latitude and longitude coordinate and color
mushroom_observer_all <- cbind(mushroom_observer_all,c(0),c(0),c(0),c(0))
#Name columns
colnames(mushroom_observer_all) <- c("observation", "specname", "north", "west", "east", "south","hemisphere","latitude","longitude","carp_color","spore_color")
#Upress rows without species name
mushroom_observer_all<-subset(mushroom_observer_all,specname!="")
#Order by species name the dataframe
mushroom_observer_all <- mushroom_observer_all[order(mushroom_observer_all[,2]), ]

## Longitude and latitude computing from coordinate gathered by php code (north, west, east, south, hemisphere)
#Calculating latitude and longitude based on location coordinate
mushroom_observer_all$latitude<-(mushroom_observer_all$north+mushroom_observer_all$south)/2
mushroom_observer_all$longitude<-(mushroom_observer_all$east+mushroom_observer_all$west)/2

#lat position 8 
#long position 9
#hemisphere 7
mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="ES"] <- -mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="ES"]
mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="SE"] <- -mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="SE"]
mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="NW"] <- -mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="NW"]
mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="WN"] <- -mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="WN"]
mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="SW"] <- -mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="SW"]
mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="WS"] <- -mushroom_observer_all$latitude[mushroom_observer_all$hemisphere=="WS"]
mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="SW"] <- -mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="SW"]
mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="NN"] <- -mushroom_observer_all$longitude[mushroom_observer_all$hemisphere=="NN"]

##Correction on mushroom observer localization
#Observation are localized by rectangular zone. Correction with a threshold of height and width. It correspond to an incertitude of 0.1° of latitude and longitude 

mushroom_observer_all<-subset(mushroom_observer_all, abs(mushroom_observer_all$west-mushroom_observer_all$east)<0.1)
mushroom_observer_all<-subset(mushroom_observer_all, abs(mushroom_observer_all$north-mushroom_observer_all$south)<0.1)

##Remove some error of mushroom user localization
#When there is no localization sometime it's logitude=-179
  mushroom_observer_all<-subset(mushroom_observer_all, mushroom_observer_all$longitude!=-179.0000)
  mushroom_observer_all<-subset(mushroom_observer_all, mushroom_observer_all$longitude!=179.0000)
  
  #Remove useless column
mushroom_observer_all$north <- NULL
mushroom_observer_all$south <- NULL
mushroom_observer_all$east <- NULL
mushroom_observer_all$west <- NULL
mushroom_observer_all$spore_color <- NULL
mushroom_observer_all$hemisphere <- NULL

#Save it
write.table(mushroom_observer_all, file = "mushroom_observer_all.csv", sep = ",")
```

#Data gathering function
```{r,echo=TRUE}
database_gathering_climate_soil<-function(dataset){
#Import map of bioclim variable
BClim = brick("data/YbrevBC_2.5.grd")
#Indicate position of different observation
points <- cbind(dataset$longitude,dataset$latitude)
#Export variable for position of observation
clim<- extract(BClim, points)
#Bind new variable to data
dataset<-cbind(dataset,clim)

##Gathering of Soil data (harmonized global database)
#Import data 
hwsd <- raster("data/HWSD_RASTER/hwsd.bil")
#Project variable into longitude latitude axis
(proj4string(hwsd) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
#Export variable for position of observation
soil1<- extract(hwsd, points)
pointcoord<-cbind(points,soil1)
colnames(pointcoord)<-c("lon","lat","hwsd")
##Use of SQLite to repair soil data
m <- dbDriver("SQLite")
con <- dbConnect(m, dbname = "data/hwsd.db")
dbWriteTable(con, name="WINDOW_BHUTAN", value=data.frame(smu_id=pointcoord[1,3]), overwrite=TRUE)
records.bhutan <- dbGetQuery(con, "select T.* from HWSD_DATA as T join WINDOW_BHUTAN as U on T.MU_GLOBAL=U.SMU_ID")
tablesoil<-records.bhutan
tablesoil<-tablesoil[0,]
for (i in 1:nrow(pointcoord)){
dbWriteTable(con, name="WINDOW_BHUTAN", value=data.frame(smu_id=pointcoord[i,3]), overwrite=TRUE)
records.bhutan <- dbGetQuery(con, "select T.* from HWSD_DATA as T join WINDOW_BHUTAN as U on T.MU_GLOBAL=U.SMU_ID")
tablesoil<-rbind(tablesoil,records.bhutan[1,])}
###Supress useless soil variable for reference
tablesoil$id<-NULL
tablesoil$mu_global<-NULL
tablesoil$mu_source1<-NULL
tablesoil$mu_source2<-NULL
tablesoil$coverage<-NULL
tablesoil$issoil<-NULL
tablesoil$seq<-NULL
tablesoil$share<-NULL
tablesoil$su_code74<-NULL
tablesoil$su_code85<-NULL
tablesoil$su_code90<-NULL
tablesoil$su_sym74<-NULL
tablesoil$su_sym85<-NULL
tablesoil$su_sym90<-NULL
#Add column of soil variable to data
dataset<-cbind(dataset,tablesoil)

#Some extra variable from harmonized global database sup
varsup = brick("varsup.grd")
points <- cbind(dataset$longitude,dataset$latitude)
varsupex<- extract(varsup, points)
dataset<-cbind(dataset,varsupex)
return(dataset)
}
```


#Save of data
```{r,echo=TRUE}
#Save it
mushroom_observer<-database_gathering_climate_soil(mushroom_observer)
#mushroom_observer_all<-database_gathering_climate_soil(mushroom_observer_all)
mushroom_observer_color<-database_gathering_climate_soil(mushroom_observer_color)
write.table(mushroom_observer, file = "mushroom_observer_amanita.csv", sep = ",")
#write.table(mushroom_observer_all, file = "mushroom_observer_all.csv", sep = ",")
write.table(mushroom_observer_color, file = "mushroom_observer_color.csv", sep = ",")
```

#Dataframe with grouped colors
```{r,echo=TRUE}
mushroom_observer_color_group <- mushroom_observer_color
levels(mushroom_observer_color_group$carp_color)<-c("grey+brown","grey+brown","red+orange+yellow","red+orange+yellow",   "white" ,"red+orange+yellow")

#Just in case
mushroom_observer_color_group <- subset(mushroom_observer_color_group, carp_color!="" )
mushroom_observer_color_group<-droplevels(mushroom_observer_color_group)
write.table(mushroom_observer_color_group, file = "mushroom_observer_color_group.csv", sep = ",")
```

#Reduction to America function
```{r,echo=TRUE}
reduction_to_usa <-function(dataset){
dataset<-dataset[dataset$latitude>25,]
dataset<-dataset[dataset$latitude<50,]
dataset<-dataset[dataset$longitude< -67,]
dataset<-dataset[dataset$longitude> -130,]
dataset<-droplevels(dataset)
return(dataset)
}
```

```{r}
#For mushroom_observer
mushroom_observer_usa<-reduction_to_usa(mushroom_observer)
write.table(mushroom_observer_usa, file = "mushroom_observer_amanita_usa.csv", sep = ",")
#For the dataframe with grouped color
mushroom_observer_color_usa<-reduction_to_usa(mushroom_observer_color)
write.table(mushroom_observer_color_usa, file = "mushroom_observer_color_usa.csv", sep = ",")
#For all
mushroom_observer_all_usa<-reduction_to_usa(mushroom_observer_all)
write.table(mushroom_observer_all_usa, file = "mushroom_observer_all_usa.csv", sep = ",")
```

#iNaturalist dataset preparation
```{r}
inaturalist_all<-inaturalist_all[inaturalist_all$coordinateUncertaintyInMeters<8800,]
inaturalist_all<-inaturalist_all[,c("id","kingdom","family","genus","scientificName","decimalLongitude","decimalLatitude")]

colnames(inaturalist_all)[colnames(inaturalist_all) == "decimalLatitude"] <-"latitude"
colnames(inaturalist_all)[colnames(inaturalist_all) == "decimalLongitude"] <-"longitude"
colnames(inaturalist_all)[colnames(inaturalist_all) == "scientificName"] <-"specname"
colnames(inaturalist_all)[colnames(inaturalist_all) == "id"] <-"observation"

#inaturalist_all<-database_gathering_climate_soil(inaturalist_all)
#All iNaturalist observation whatever the organism is
write.csv(inaturalist_all ,file = "inaturalist_all.csv")
#Reduction to usa
inaturalist_all_usa<-reduction_to_usa(inaturalist_all)
write.csv(inaturalist_all_usa ,file = "inaturalist_all_usa.csv")

#Subset observation to amanita genus
inaturalist_amanita<-subset(inaturalist_all,genus=="Amanita")
write.csv(inaturalist_amanita ,file = "inaturalist_amanita.csv")

#Reduction to USA
inaturalist_amanita_usa<-reduction_to_usa(inaturalist_amanita)
write.csv(inaturalist_amanita_usa ,file = "inaturalist_amanita_usa.csv")

#Subset observation to plantae kingdom
inaturalist_plantae<-subset(inaturalist_all,kingdom=="Plantae")
write.csv(inaturalist_plantae ,file = "inaturalist_plantae.csv")

```

#Gbif dataset preparation
```{r}
gbif_amanita<-gbif_amanita[gbif_amanita$coordinateUncertaintyInMeters<8800,]
gbif_amanita<-gbif_amanita[,c("kingdom","family","genus","scientificName","lon","lat")]
colnames(gbif_amanita)[colnames(gbif_amanita) == "lat"] <-"latitude"
colnames(gbif_amanita)[colnames(gbif_amanita) == "lon"] <-"longitude"
colnames(gbif_amanita)[colnames(gbif_amanita) == "scientificName"] <-"specname"
write.csv(gbif_amanita,file = "gbif_amanita.csv")
```


#Dictionnary of Amanita species
```{r}
spec_dict<-read.csv("spec_dict.csv",header=FALSE)
rownames(spec_dict)<-c()
colnames(spec_dict)<-c()
spec_dict<-t(spec_dict)
spec_dict<-spec_dict[,2]
spec_dict<-gsub(pattern="(</i>)","" ,x=spec_dict)
spec_dict<-gsub(pattern="(<i>)","" ,x=spec_dict)
spec_dict<-sub("\\s+$", "", spec_dict)
spec_dict<-paste("Amanita",spec_dict)
write.table(spec_dict ,file = "amanita_species_name.csv",sep=",")
```





